---
title: "SEDA EDA"
author: "Andy Cooke"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare Data

```{r}
library(dplyr)
library(haven)
library(ggmap)
library(data.table)
library(mapproj)
library(GGally) 
```


# is grade nested in year or 1:1?

```{r}
files <- list.files(path = "./SEDA/", pattern = "*.dta")
datasets <- list()
for(i in 1:length(files)){
  datasets <- append(datasets, list(read_dta(paste0("SEDA/", files[i]))))
}
files <- chartr(" ", "_", files)
files <- sub(".dta", "", files)
files <- chartr(".", "_", files)
names(datasets) <- files
```

```{r}
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

remove_y_cols <- function(data){
  tmp <- colnames(data)
  data <- data %>% select(tmp[substrRight(tmp, 2) != '.y'])
  tmp <- colnames(data)
  tmp <- gsub(".x", "", tmp)
  colnames(data) <- tmp
  return(data)
}
```

```{r}
# still duplicating columns after removing .y??
county_joins <- c('sedacounty', 'subcat', 'subgroup', 'fips', 'stateabb', 'gradecenter', 'gap', 'sedacountyname')
seda_counties <- datasets$seda_county_pool_cs_4_1
seda_counties <- merge(seda_counties, datasets$seda_county_pool_gcs_4_1, all =T, by = county_joins)
seda_counties <- merge(seda_counties, datasets$seda_county_poolsub_cs_4_1, all =T, by = county_joins)
seda_counties <- merge(seda_counties, datasets$seda_county_poolsub_gcs_4_1, all =T, by = county_joins)

seda_counties <- remove_y_cols(seda_counties)

seda_counties_all <- seda_counties %>% filter(subcat == 'all')
seda_counties_race <- seda_counties %>% filter(subcat == 'race')
seda_counties_gender <- seda_counties %>% filter(subcat == 'gender')
seda_counties_ecd <- seda_counties %>% filter(subcat == 'ecd')

long_joins <- c('sedacounty', 'year', 'grade', 'fips')
seda_long <- datasets$seda_covariates_test_scores_marijuna
seda_long <- merge(seda_long, datasets$seda_cov_county_long_4_1, all=T, by = long_joins)%>%
  distinct(sedacounty, year, grade, .keep_all = TRUE)
seda_long <- remove_y_cols(seda_long)
seda_long <- merge(seda_long, datasets$seda_cov_county_poolyr_4_1, all = T, by = c('year', 'sedacounty'))
seda_long <- remove_y_cols(seda_long)
dupes <- colnames(seda_long)
dupes <- dupes[duplicated(dupes)]
seda_long <- seda_long %>% select(-dupes)
```

# Descriptive

```{r}
seda_long_sort <- seda_long %>%
  arrange(sedacounty, grade, year)
seda_long_sort
```

Year is nested in grade which is nested in county.

```{r}
county_pair_cols <- c('tot_asmts', 'cellcount', 'mn_asmts', 'cs_mn_avg_ol', 
                      'cs_mn_coh_ol', 'gcs_mn_avg_ol') # add more
ggpairs(subset(seda_counties_all, select = county_pair_cols))
```

```{r}
corr_county <- cor(na.omit(seda_counties_all[,9:ncol(seda_counties_all)]))
corr_county
```

There is high correlation between variables with similar suffices and gcs to cs counterparts.

```{r}
long_pair_cols <- c("urban", "perfl", "totenrl", "lninc50asn", "povertyall", "unempall",
                    "sesblk", "snapall")
ggpairs(subset(seda_long, select = long_pair_cols))
```

```{r}
corr_long <- cor(na.omit(subset(seda_long, select = names(seda_long[, sapply(seda_long, is.numeric)]))))
corr_long
```

Some columns have 100% correlation, some have only one value. These can be cleaned for further analysis.

# Maps

```{r}
county_fix <- function(col){
  # make it a loop
  col = tolower(col)
  col = sub(" county", "", col)
  col = sub(" census area", "", col)
  col = sub(" parish", "", col)
  col = sub(" municipio", "", col)
  col = sub(" borogh", "", col)
  col = sub(" city", "", col)
  col = sub("'", "", col)
  col = sub(" municipality", "", col)
  col = sub("dekalb", "de kalb", col)
  col = sub("dewitt", "de witt", col)
  col = sub("desoto", "de soto", col)
  col = sub("dupage", "du page", col)
  col = sub("lasalle", "la salle", col)
  col = sub("o'brien", "obrien", col)
  col = sub("oglala lakota", "oglala dakota", col)
  col = sub("st. ", "st ", col)
  return(col)
}
```


```{r}
neglog <- function(col){
  c_sign <- if_else(col >= 0, 1, -1)
  av <- abs(col)
  lcol <- log(av)
  final <- lcol * c_sign
  final <- if_else(final == -Inf, 0, final)
  return(final)
}
```

```{r}
# need one row per county or one row per county per facet!
county_map_plot <- function(data, target, county_col = 'sedacountyname', 
                            state_abb_col = "stateabb", log = F, facet_col = 'None'){
  data[,'county_col'] <- data[,county_col]
  data$c_fix <- county_fix(data$county_col)
  data[,'state_abb_col'] <- data[,state_abb_col]
  data[,'target'] <- data[,target]
  if(facet_col != 'None'){
    data[,'facet_col'] <- data[,facet_col]
  }
  
  counties <- map_data("county")
  states <- data.frame(name = tolower(state.name), abb = state.abb)
  counties <- merge(counties, states, all.x = TRUE, by.x = "region", by.y = "name")
  
  if(log == T){
    data$target <- neglog(data$target)
  }
  data[,target] = data$target
  
  data <- merge(counties, data, all.x=TRUE, by.y = c("c_fix", "state_abb_col"), by.x = c('subregion', 'abb')) %>%
    arrange(order)
  
  p <- ggplot(data, aes(x=long, y=lat, group=group, fill=target)) + 
    ggtitle(paste0(target, " by county")) + 
    geom_polygon()+
    coord_map()
  if(facet_col != 'None'){
    p <- p + facet_wrap(~facet_col)
  }
  print(p)
}
```


```{r}
county_map_cols <- c(county_pair_cols)
log_val <- c(T, F, T, F, F, F)

for(i in 1:length(county_map_cols)){
  county_map_plot(seda_counties_all, county_map_cols[i], log = log_val[i])
}
```

```{r}
for(i in 1:length(county_map_cols)){
  county_map_plot(seda_counties_race, county_map_cols[i], log = log_val[i], facet_col = 'subgroup')
}
```

```{r}
for(i in 1:length(county_map_cols)){
  county_map_plot(seda_counties_gender, county_map_cols[i], log = log_val[i], facet_col = 'subgroup')
}
```

```{r}
for(i in 1:length(county_map_cols)){
  county_map_plot(seda_counties_ecd, county_map_cols[i], log = log_val[i], facet_col = 'subgroup')
}
```


# By Grade

## Univariate
```{r}
grade_plot <- function(data, target){
  data[,'target'] <- data[,target]
  
  data <- data %>% 
    group_by(grade, year) %>% 
    summarise(target = mean(target, na.rm = T))
  
  p <- ggplot(data, aes(x=year, y=target,color = as.factor(grade))) +
    geom_line() + 
    ggtitle(paste0(target, " by grade over time"))
  print(p)
}
```


```{r}
long_targets <- c(long_pair_cols)
for(i in long_targets){
  grade_plot(seda_long, i)
}
```


## Followup: over time by cohort instead of moving grade








